Interisland movements and oceanic swimming of woodland caribou in Newfoundland

### 2020-05-11 paper/manuscript.Rmd
P4

Caribou (*Rangifer tarandus*) are exceptional swimmers. Ample evidence exists
that caribou swim in streams, rivers, and lakes during migration [@Leblond2016],
to avoid predators [@Bergerud1985], and access islands during calving
[@Bergerud1990]. Even for adept swimmers like caribou, the energetic expenditure
of swimming for quadrupedal mammals is significantly higher than walking or
running, and drowning is also possible [@Miller1985]. Despite
the abundance of coastal and island caribou herds, only a handful of examples
have documented caribou swimming in the ocean [e.g. @Miller1995; @Miller2002;
@Jeffery2007; @Ricca2012].
Forage limitation and over-grazing is a major concern for caribou populations:
it can reduce female reproductive success [@Schaefer2016] and thus could lead to
movement between islands. Forage scarcity has been proposed as a potential
reason for caribou moving between arctic islands on the sea-ice in winter
[@Miller1977]. Caribou were introduced to Fogo Island (n = 26 animals between
1964–67) and Change Island (n = 5 animals in 1964) from Newfoundland as part of
a series of translocations and introductions throughout the province (Bergerud &
Mercer 1989). It is unknown when caribou began colonizing nearby islands, but
given that some are small, over-grazing by newly arrived caribou may rapidly
deplete forage on an annual cycle [@Bergerud1990]. Although unconfirmed, it is
also possible that Fogo Island has experienced forage depletion since the
introduction of caribou in the 1960s. More likely, however, is that fine-scale
competition for foraging resources drives caribou to swim to new islands.

@@

Caribou may swim between islands in the ocean for similar
reasons they swim in freshwater [@Leblond2016]. We propose that one such
explanation, the forage limitation hypothesis, is a likely explanation for
oceanic swimming for caribou that live on islands in the Fogo Island
archipelago.

>>>>

Caribou (*Rangifer tarandus*) are exceptional swimmers. Ample evidence exists
that caribou swim in streams, rivers, and lakes during migration [@Leblond2016],
to avoid predators [@Bergerud1985], and access islands during calving
[@Bergerud1990]. Even for adept swimmers like caribou, the energetic expenditure
of swimming for quadrupedal mammals is significantly higher than walking or
running, and drowning is also possible [@Miller1985]. Despite
the abundance of coastal and island caribou herds, only a handful of examples
have documented caribou swimming in the ocean [e.g. @Miller1995; @Miller2002;
@Jeffery2007; @Ricca2012].

Caribou may swim between islands in the ocean for similar
reasons they swim in freshwater [@Leblond2016], mainly concerning predator avoidance or movement between habitats. We propose that one such
explanation, the forage limitation hypothesis, is the most likely explanation for
oceanic swimming for caribou that live on islands in the Fogo Island
archipelago.

Forage limitation and over-grazing is a major concern for caribou populations:
it can reduce female reproductive success [@Schaefer2016] and thus could lead to
movement between islands. Forage scarcity has been proposed as a potential
reason for caribou moving between arctic islands on the sea-ice in winter
[@Miller1977]. Caribou were introduced to Fogo Island (n = 26 animals between
1964–67) and Change Island (n = 5 animals in 1964) from Newfoundland as part of
a series of translocations and introductions throughout the province (Bergerud &
Mercer 1989). It is unknown when caribou began colonizing nearby islands, but
given that some are small, over-grazing by newly arrived caribou may rapidly
deplete forage on an annual cycle [@Bergerud1990]. Although unconfirmed, it is
also possible that Fogo Island has experienced forage depletion since the
introduction of caribou in the 1960s. More likely, however, is that fine-scale
competition for foraging resources drives caribou to swim to new islands.

@@
